---
title: "cancer viz"
author: "Connor Flynn"
date: "7/21/2024"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(plotly)
library(viridis)
library(ggstream)
library(writexl)
library(forcats)
library(viridis)
library(sunburstR)
```

Cancer in New Caledonia
- With 162 registered cases, the ASR for all cancers combined was 142.2 (range, 104.9 to 193.3) cases per million children per year and was not different from what was estimated in France (156.6
- 1985 -> 1999 Thyroid cancer incidence in New Caledonia is the highest reported in the world and is approximately 10-fold higher than in most developed countries.
- New Caledonia not isolated in Our World in Data



```{r}
nc_cancer_data <- read_csv(here("data/cas_cancer_nc.csv"))
```

```{r}
# Adding the broader category column
nc_cancer_data <- nc_cancer_data %>%
  mutate(category = case_when(
    `ICD-10` %in% c("C00", "C01", "C02", "C03", "C04", "C05", "C06", "C07", "C08", "C09", "C10", "C11", "C12", "C13", "C14") ~ "Malignant neoplasms of lip, oral cavity and pharynx",
    `ICD-10` %in% c("C15", "C16", "C17", "C18", "C19", "C20", "C21", "C22", "C23", "C24", "C25", "C26") ~ "Malignant neoplasms of digestive organs",
    `ICD-10` %in% c("C30", "C31", "C32", "C33", "C34", "C37", "C38", "C39") ~ "Malignant neoplasms of respiratory and intrathoracic organs",
    `ICD-10` %in% c("C40", "C41") ~ "Malignant neoplasms of bone and articular cartilage",
    `ICD-10` %in% c("C43", "C44") ~ "Melanoma and other malignant neoplasms of skin",
    `ICD-10` %in% c("C45") ~ "Mesothelial and soft tissue neoplasms",
    `ICD-10` %in% c("C46") ~ "Kaposi's sarcoma",
    `ICD-10` %in% c("C47", "C48", "C49") ~ "Malignant neoplasms of mesothelial and soft tissue",
    `ICD-10` %in% c("C50") ~ "Malignant neoplasms of breast",
    `ICD-10` %in% c("C51", "C52", "C53", "C54", "C55", "C56", "C57", "C58") ~ "Malignant neoplasms of female genital organs",
    `ICD-10` %in% c("C60", "C61", "C62", "C63") ~ "Malignant neoplasms of male genital organs",
    `ICD-10` %in% c("C64", "C65", "C66", "C67", "C68") ~ "Malignant neoplasms of urinary tract",
    `ICD-10` %in% c("C69", "C70", "C71", "C72") ~ "Malignant neoplasms of eye, brain, and other parts of the central nervous system",
    `ICD-10` %in% c("C73", "C74", "C75") ~ "Malignant neoplasms of endocrine glands",
    `ICD-10` %in% c("C76", "C77", "C78", "C79", "C80") ~ "Malignant neoplasms of ill-defined, secondary and unspecified sites",
    `ICD-10` %in% c("C81", "C82", "C83", "C84", "C85", "C86", "C88") ~ "Malignant neoplasms of lymphoid, hematopoietic and related tissue",
    `ICD-10` %in% c("C90", "C91", "C92", "C93", "C94", "C95", "C96") ~ "Malignant neoplasms of lymphoid, hematopoietic and related tissue",
    TRUE ~ "Other"
  ))
```

```{r}
#nc_cancer_data$counter_column = 1
```

```{r}
#write_csv(nc_cancer_data, here("data/nc_cancer_data_categories.csv"))
```

```{r}
#write_xlsx(nc_cancer_data, here("data/nc_cancer_data_categories.xlsx"))
```


```{r}
nc_cancer_data_grouped_category <- nc_cancer_data %>% 
  group_by(sex, diagnosis_year, category) %>% 
  count()
```


```{r}
# Create the plot

ggplot(data = nc_cancer_data_grouped_category, aes(x = diagnosis_year, 
                                                  y = n,
                                                  fill = category)) +
  geom_area(color = "white") +
  facet_wrap(~sex, nrow=2) +
  theme_bw() +
  scale_fill_viridis_d() +
  labs(x = "Year", y = "Count", fill = "Category")
```




```{r}
a <- ggplot(data = nc_cancer_data_grouped_category, aes(x = diagnosis_year, 
                                                  y = n,
                                                  color = category)) +
  geom_line() +
  facet_wrap(~sex, nrow=2) +
  theme_bw() +
  scale_color_viridis_d() +
  labs(x = "Year", y = "Count", fill = "Category")

ggplotly(a)
```


Petal Plot

```{r}
nc_cancer_data_petal <- nc_cancer_data %>% 
  group_by(age_class, category) %>% 
  count()
```

```{r}
# Reorder the age_class factor levels
nc_cancer_data_petal$age_class <- factor(
  nc_cancer_data_petal$age_class,
  levels = c("moins de 20 ans", "20-29 ans", "30-39 ans", "40-49 ans", "50-59 ans", "60-69 ans", "70-79 ans", "80 ans et +")
)

# Verify the factor levels
print(levels(nc_cancer_data_petal$age_class))
```



```{r, fig.width=10, fig.height=8}
# Define the Viridis color palette
viridis_colors <- viridis_pal(option = "B")(length(unique(nc_cancer_data_petal$category)))

# Create the plot
fig <- plot_ly() %>%
  add_trace(
    data = nc_cancer_data_petal,
    r = ~n, # radius 
    theta = ~age_class, # angle
    type = "barpolar", 
    color = ~category,
    colors = viridis_colors,
    hovertemplate = paste('n: %{r}',
                          '<br>age_class: %{theta}<br>')) %>%  
  layout(
    showlegend = TRUE, # keep legend
    legend = list(
      font = list(size = 10) # make the legend text smaller
    ),
     polar = list(
      bgcolor = "#1f1f1f",
      angularaxis = list(
        rotation = 90,
        direction = 'clockwise',
        categoryorder = "array", # explicit ordering of categories
        categoryarray = c("moins de 20 ans", "20-29 ans", "30-39 ans", "40-49 ans", "50-59 ans", "60-69 ans", "70-79 ans", "80 ans et +"),
        period = 8,
        tickcolor = "#ffffff",
        gridcolor = "#444444"),
      radialaxis = list(
        tickcolor = "#ffffff",
        gridcolor = "#444444")),
    paper_bgcolor = "#1f1f1f",
    plot_bgcolor = "#1f1f1f",
    font = list(color = "#ffffff"),
    title = list(
      text = "Cancer Instances By Age Class, New Caledonia",
      font = list(size = 16, color = 'white')
    )
  )

fig

```

Animate?


Suburst plot



```{r}
nc_cancer_data_sunburst <- nc_cancer_data %>% 
  group_by(long_label_disease, category) %>% 
  count(name = "total_cases")
```

```{r}
nc_cancer_data_sunburst <- nc_cancer_data_sunburst %>%
     mutate(category = str_replace(category, "Malignant neoplasms of ill-defined, secondary and unspecified sites", 
                                  "Malignant neoplasms of illdefined, secondary and unspecified sites"))
            
nc_cancer_data_sunburst <- nc_cancer_data_sunburst %>%
  mutate(long_label_disease = str_replace(long_label_disease, 
                                 "Tumeur maligne de la jonction recto-sigmoïdienne", 
                                 "Tumeur maligne de la jonction rectosigmoïdienne"))

```


```{r}
nc_cancer_data_sunburst <- nc_cancer_data_sunburst %>% 
  mutate(path = paste(category, long_label_disease, sep="-")) %>% 
  ungroup() %>%
  dplyr::select(path, total_cases)
```


```{r}
p <- sunburst(nc_cancer_data_sunburst, legend=FALSE, count = TRUE)
p
```



